---
title: "Billigence1"
output:
  pdf_document: default
  html_document: default
date: "2024-03-17"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## R Markdown


```{r }
# install.packages("dyplr")
# install.packages("mlbench")
library(dplyr)
library(tidyverse)
library(lubridate)
library(tidyr)
library(skimr)
library(ggplot2)
library(mlbench)
library(gmodels)
library(class)
library(knitr)


```


```{r}

#reading the client dataset
client_df<-read.csv("Data.csv")

#checking first 6 entries 
head(client_df)

#smmarising
summary(client_df)



```
```{r}

#renaming the column "y" to "State" 
#which describes whether the customer said 'yes' or 'no' to the deposit

client_df <- client_df %>% 
       rename("state" = "y")
head(client_df)




```

Here the goal is to determine whether the customer will say yes to deposit. Since the dataset is large I will be focussing majorily on customers who said yes. And try to find any relation or pattern among the customers. So in future the bank could focus on marketing among such customers.

```{r, 'filter'}
# 
# #filtering 
# 
df_state_yes<-filter(client_df,state=="yes")
head(df_state_yes)

# 
# # Calculate count by age and state
# client_count <- client_df %>%
#   group_by(age, state) %>%
#   summarize(Count = n())
# 
# 
# # Define custom color palette
# custom_palette <- c("blue", "red")
# 
# # Apply custom theme
# theme_set(theme_minimal(base_size = 12) +
#             theme(axis.title.x = element_text(size = 14),
#                   axis.title.y = element_text(size = 14),
#                   axis.text.x = element_text(size = 10),
#                   axis.text.y = element_text(size = 10),
#                   plot.title = element_text(hjust = 0.5, size = 16),
#                   legend.position = "bottom"))
# 
# # Plot
# ggplot(data = client_count, aes(x = age, y = Count, color = state, group = state)) +
#   geom_line(size = 1.2) +
#   scale_x_continuous(breaks = seq(20, 70, by = 5)) +
#   scale_color_manual(values = custom_palette, name = "State",
#                      labels = c("Yes", "No")) +
#   labs(title = "Relationship between Age and Status",
#        x = "Age",
#        y = "Count",
#        color = "State")


```
morecritically evaluate the policies and think i through thats why max rejectios and acceptions ijn this area
According to the graph , majority of the customers who said yes to deposit is of age between 25 and 65.

```{r}

unique(client_df$state)

#remove empty string values form state column
client_df <- subset(client_df, state != " ")

unique(client_df$state)



ggplot(data = client_df, aes(x = state, y = duration, fill = state, colour = state)) + 
  geom_bar(stat = "identity") 
 



```

The relationship between the duration of call and status for all customers(yes and no) clearly shows no connection between the duration of call and status.



```{r}


#Relationship between Campaign and Status
ggplot(data=client_df, aes(x = state,y= campaign , colour = state)) + 
  geom_bar(stat = "identity")




```
This graph shows no positive relation is there between the number of times contacted during this campaign and status.





```{r}

#Relationship between Marital status and Age

ggplot(df_state_yes)+
  geom_point(mapping = aes(x=marital,y=age) ,color="blue")



```
The graph shows that

Single are more likely to say yes to deposit around age 20 to 55
Married people prefer deposits aged around 25 to 85
Divorced people prefer deposits aged between late 25 to 65


```{r}


#Relationship between Occupation and Status
ggplot(data = client_df) +
  geom_bar(mapping = aes(x = job, fill = state), color = "blue") +
  theme(axis.text.x = element_text(angle = 45), legend.position = "bottom") +
  labs(x = "Occupation", title = "Relationship between Occupation and Status")
```

The graph shows the occupation of customers who said yes to deposits. According to the data , the customers whose occupation is management related are those who highly prefer deposits, followed by technician, followed by blue collar etc.


```{r}


#Relationship between Education and Status

ggplot(data=client_df)+
  geom_bar(mapping = aes(x = education,fill=state))+ 
  theme(axis.text.x = element_text(angle = 45)) +
  labs(title = "Relationship between Education and Status")




```
According to the dataset, majority of customers who said yes to deposits are secondary level educated, followed by teritiary.


```{r}

#Relationship between Loan and Status

ggplot(data=client_df)+
  geom_bar(mapping = aes(x = loan,fill=state))+ 
  labs(title = "Relationship between Loan and Status")




```

According to the dataset, majority of customers who said yes to deposits doesnt have loan.

```{r}


day_hist <- ggplot(df_state_yes, aes(day))
day_hist + geom_area(stat = "bin", color = "black",fill = "slateblue1", alpha = 0.5) + theme(text = element_text(size=30))



```

```{r}

conversionsAgeMarital <- df_state_yes %>%
  group_by(AgeGroup = cut(age, breaks = seq(20, 70, by = 10)),
           Marital = marital) %>%
  summarize(Count = n(), 
            NumConversions = sum(ifelse(state == "yes", 1, 0))) %>%
  mutate(TotalCount = sum(Count),
         ConversionRate = NumConversions / TotalCount * 100)

# Print the intermediate steps to debug
print(conversionsAgeMarital)

kable(conversionsAgeMarital)



#rename the last groups
conversionsAgeMarital$AgeGroup <- as.character(conversionsAgeMarital$AgeGroup)
conversionsAgeMarital$AgeGroup[is.na(conversionsAgeMarital$AgeGroup)] <- "70+"

#visualizing conversions by age group and marital status
ggplot(conversionsAgeMarital, aes(x=AgeGroup, y=ConversionRate, fill=Marital)) +
  geom_bar(width=0.5, stat = "identity") +
  labs(title="Conversion Rates by Age Group and Marital Status") +
  theme( legend.position = 'bottom')




# Sort the data frame by ConversionRate in descending order
conversionsAgeMarital <- conversionsAgeMarital[order(-conversionsAgeMarital$ConversionRate), ]

# Print the sorted data frame
print(conversionsAgeMarital)

# Rename the last groups
conversionsAgeMarital$AgeGroup <- as.character(conversionsAgeMarital$AgeGroup)
conversionsAgeMarital$AgeGroup[is.na(conversionsAgeMarital$AgeGroup)] <- "70+"

# Tabulate the sorted data
kable(conversionsAgeMarital)
write.csv(conversionsAgeMarital, "conversions_data.csv", row.names = TRUE)


```
```{r}




client_df$month <- factor(client_df$month, levels = c("jan", "feb", "mar", "apr", "may", "jun", "jul", "aug", "sep", "oct", "nov", "dec")) # Sort by chronological month order

month_bar <- ggplot(client_df, aes(month, fill = state)) +
  geom_bar(color = "black") + 
  theme(text = element_text(size = 30), axis.text.x = element_text(angle = 90, vjust = 0.5, hjust = 1, size = 30))

month_bar





```

```{r}


df_state_yes$month <- factor(df_state_yes$month,levels = c("jan", "feb", "mar", "apr","may", "jun", "jul", "aug","sep", "oct", "nov", "dec")) #sort by chronological month order
month_bar <- ggplot(df_state_yes, aes(month))
month_bar + geom_bar(color = "black",fill = "gold") +  theme(text = element_text(size=30), axis.text.x=element_text(angle = 90, vjust = 0.5, hjust=1,size=30))




```

```{r}


df_state_no <- filter(client_df,state=="no")
head(df_state_no)


df_state_no$month <- factor(df_state_no$month,levels = c("jan", "feb", "mar", "apr","may", "jun", "jul", "aug","sep", "oct", "nov", "dec")) #sort by chronological month order
month_bar <- ggplot(df_state_no, aes(month))
month_bar + geom_bar(color = "black",fill = "gold") +  theme(text = element_text(size=30), axis.text.x=element_text(angle = 90, vjust = 0.5, hjust=1,size=30))





```





Recommendations
According to the insights gained from this analysis, bank should put more focus on following categories in future campaigns:

Aged between 25 to 85
Primary focus on married , followed by young single then divorcee
Education level: secondary and higher
Occupation: Management followed by technician,then blue collar , admin.

